---
title: "Assignment 1"
author: 
  - Albert Kaczmarek, h52201551
  - Maciej Kilijański, h12018188
  - Balázs Kiss, h12314734
output: 
  github_document:
    toc: TRUE
  pdf_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Setup: 

Loading necessary libraries.
```{r,message=FALSE, warning=FALSE}
library("tidyverse")
library("ggplot2")
library("readr")
library("lubridate")
library("dplyr")
library("ggplot2")
```

## Part 1: Data Loading and Inspection

### 1. Load the Data
```{r,message=FALSE}
customer_sales <- read_csv("customer_sales.csv")
customer_info <- read_csv("customer_info.csv")
product_info <- read_csv("product_info.csv")
supplier_info <- read_csv("supplier_info.csv")
```

### 2. Inspect the Data

```{r}
glimpse(customer_sales)
summary(customer_sales)

glimpse(customer_info)
summary(customer_info)

glimpse(product_info)
summary(product_info)

glimpse(supplier_info)
summary(supplier_info)
```

## Part 2: Data Cleaning and Tidying

### 3. Handle Missing Values in customer_sales

Identify rows with missing Quantity, UnitPrice, or TotalPrice:
```{r}
customer_sales |> 
  filter(is.na(Quantity) | is.na(UnitPrice) | is.na(TotalPrice))
```

Identifying rows with available Quantity, UnitPrice, but no TotalPrice

```{r}
customer_sales |> 
  filter(!is.na(Quantity) & !is.na(UnitPrice) & is.na(TotalPrice))
```
> There are 5 such values

Correcting these values:, using `mutate()`. If `TotalPrice` is missing, it will be calculated by multiplicating $Quantity\cdot UnitPrice$.
```{r}
customer_sales <- customer_sales |> 
  mutate(TotalPrice = ifelse(!is.na(Quantity) & !is.na(UnitPrice & is.na(TotalPrice)), 
                             Quantity * UnitPrice, 
                             TotalPrice))
```
Filtering rows where `Quantity` and `TotalPrice` are available, but `UnitPrice` is missing:
```{r}
customer_sales |> 
  filter(!is.na(Quantity) & is.na(UnitPrice) & !is.na(TotalPrice))
```
> There are no such rows

We do not need to fill in missing `UnitPrice`.

Filtering rows where Unit Price and TotalPrice are available, Quantity is missing:
```{r}
customer_sales |> 
  filter(is.na(Quantity) & !is.na(UnitPrice) & !is.na(TotalPrice))
```
> There are no such rows

We do not need to fill in missing `Quantity`.

> The rows with remaining NA values in 2 or 3 of the 3 columns are removed since their values cannot be inferred:
```{r}
customer_sales <- customer_sales |> 
  filter(!(is.na(Quantity) | is.na(UnitPrice) | is.na(TotalPrice)))
```

Only 70 rows are kept from the original 80:

* 10 rows have 2 out of 3 crucial values missing, making them impossible to inferr.


### 4. Standardize Date Formats

Substituting values which parse_date_time() doesn't recognize:
```{r}
customer_sales$Date <- gsub("Jän", "Jan", customer_sales$Date)
```
After the values have been cleaned, we can parse the date column:
```{r}
customer_sales$Date <- parse_date_time(
  customer_sales$Date,
  orders = c("ymd", "dmy", "mdy", "Ymd", "dmY", "mdY", "dMY"),
  locale = "de_DE" # This helps parse German month abbreviations
)
```

Let's see the result (`head(,3)` will display 3 top rows):
```{r}
head(customer_sales,3)
```

Correcting the dates which weren't interpreted for 2023 but should've been
e.g. `30-03-23` was parsed as `2030-03-23` instead of `2023-03-30`.
We perform this in several steps:

> Splitting the data frame into 2 - 2023 and non-2023 (wrong) dates
1. Filter observations where Date is for 2023 or not available
```{r}
df_2023 <- customer_sales |> 
  filter(year(Date) == 2023 | is.na(Date)) |> print()
```
2. Filter observations where Date is NOT for 2023
```{r}
df_filtered <- customer_sales |> 
  filter(year(Date) != 2023)  |> print()
```
3. Converting dates from YYYY-MM-DD to YY-MM-DD
```{r}
df_filtered <- df_filtered |> 
  mutate(Date = format(Date, "%y-%m-%d")) |> print()
```
4. Converting all non-2023 values to 2023 (Switching YY with DD)
```{r}
df_filtered <- df_filtered |> 
  mutate(Date = dmy(Date)) |> print()
```
5. Converting them back to standard format (YYYY-MM-DD)
6. Combining the 2 data frames back together
```{r}
customer_sales <- bind_rows(df_2023, df_filtered) |> print()
```
7. Checking if all dates are for 2023
```{r}
customer_sales |>
  filter(year(Date) != 2023)
```

As we can see, the returned tibble is empty, meaning all dates are correct.
We can remove the 2 separate parts as we don't need them anymore
```{r}
rm(df_2023, df_filtered)
```
### 5. Correct data types

Converting OrderID, Quantity, UnitPrice and TotalPrice columns from dbl to integer.
```{r}
customer_sales$OrderID <- as.integer(customer_sales$OrderID)
customer_sales$Date <- as.Date(customer_sales$Date)
customer_sales$Quantity <- as.integer(customer_sales$Quantity)
customer_sales$UnitPrice <- as.integer(customer_sales$UnitPrice)
customer_sales$TotalPrice <- as.integer(customer_sales$TotalPrice)
```
Let's see the result:
```{r}
head(customer_sales,3)
```
All formats are correct

### 6. Resolve inconsistencies

Correct any discrepancies found + Correcting negative values
```{r}
customer_sales$UnitPrice <- abs(customer_sales$UnitPrice)
customer_sales$TotalPrice <- abs(customer_sales$TotalPrice)
```
Verify that TotalPrice equals $Quantity \cdot UnitPrice$ for all rows
```{r}
all(customer_sales$TotalPrice == customer_sales$Quantity * customer_sales$UnitPrice)
```

> All values correspond with each other


## Part 3: Data Integration with joins

### 7. Left join
Perform a `left_join()` of `customer_sales` with `customer_info` on `CustomerID`
```{R}
customer_sales_enriched <- left_join(customer_sales, customer_info, by = "CustomerID")
```

Identify how many rows have missing customer information after the join
```{r}
customer_sales_enriched |> filter(is.na(Name))
```
> There are 3 observations with missing customer information

### 8. Inner join

Perform a `inner_join()` of `customer_sales_enriched` with `product_info` on `ProductID`
```{r}
customer_sales_enriched <- inner_join(
  customer_sales_enriched, product_info, by = "ProductID"
  )
```

Determine the number of rows in the resulting dataset
```{r}
glimpse(customer_sales_enriched)
```

> 67 rows instead of 70 in the previous dataset

### 9. Right join

Perform a `right_join()` of `product_info` with `customer_sales` on `ProductID`
```{r}
product_info_enriched <- right_join(product_info, customer_sales, by = "ProductID")
```

Identify products that have no sales records
```{r}
product_info_enriched |> 
  filter(is.na(OrderID))
```

> There is no such product

### 10. Full Join

Perform a `full_join()` of `customer_info` with `customer_sales` on `CustomerID`
```{r}
customer_info_enriched <- full_join(customer_info, customer_sales, by = "CustomerID")
```

Identify customers who have not made any purchases
```{r}
customer_info_enriched |> 
  filter(is.na(OrderID))
```

> 1 customer hasn't made any purchases: Oscar Jones


## Part 4: Data Manipulation with dplyr

### 11. Create a Profit and an AgeGroup column

```{r}
customer_sales_enriched <- customer_sales_enriched |> 
  mutate(Profit = TotalPrice * 0.25,
         AgeGroup = case_when(
           Age < 30 ~ "Under 30",
           Age >= 30 & Age <= 40 ~ "30-40",
           Age > 40 ~ "Over 40"
         ))
```

### 12. Filter and select data

Assuming we want to get info about the customer region, not the sales region, we will use `Region.y`
```{r}
customer_sales_enriched |> 
  filter(Category == "Electronics") |> 
  select(OrderID,
         Date,
         CustomerID,
         Name,
         Product,
         Quantity,
         UnitPrice,
         TotalPrice,
         Profit,
         Region.y,
         AgeGroup)
```

### 13. Arrange Data

Arranging descending by `TotalPrice`:
```{r} 
customer_sales_enriched |>
  arrange(desc(TotalPrice))
```
## Part 5: Data Aggregation

### 14. Group and Summarize

Grouping by `Region.x` (sales region) and `Category`. Then creating a summary of `TotalQuantity`, `TotalPrice` and `AverageProfit`.
```{r}
customer_sales_enriched |> 
  group_by(Region.x, Category) |>
  summarise(TotalQuantity = sum(Quantity),
            TotalPrice = sum(TotalPrice),
            AverageProfit = mean(Profit), .groups = 'keep') |>
            arrange(Region.x, Category)
```

### 15. Compute cumulative metrics

Computing `TotalPrice` per `Region.x`, arranged descending by `Region.x`
```{R}
customer_sales_enriched |> 
  group_by(Region.x) |> 
  summarise(TotalPrice = sum(TotalPrice)) |>
  arrange(desc(Region.x))
```
## Part 6: Data Tidying with tidyr
### 16. Reshape the data

Creating a wide-format table showing TotalPrice for each Product across different Regions
```{r}
wide <- customer_sales_enriched |>
  group_by(ProductID, Region.x) |>
  summarise(TotalPrice = sum(TotalPrice)) |>
  pivot_wider(names_from = Region.x, values_from = TotalPrice)
wide
```

> New columns called afrer `Region.x` values. Less rows.

Converting it back to long format
```{r}
long <- wide |> 
  pivot_longer(
    cols = c("East", "North", "South", "West"),
    names_to = 'Region',
    values_to = 'TotalPrice'
  ) |> 
  filter(!is.na(TotalPrice))
long
```
> Less columns, back to 48 rows

## Part 7: Data Visualization with ggplot2

### 17. Create a bar chart

Plotting total price for each category on bar chart
```{r}
customer_sales_enriched |> 
  group_by(Category) |> 
  summarise(TotalPrice = sum(TotalPrice)) |> 
  ggplot(mapping = aes(x = Category)) + 
    geom_bar(aes(weight = TotalPrice))
```


### 18. Creating a time series plot

```{r}
customer_sales_enriched |> 
  ggplot(mapping = aes(x = Date, y = TotalPrice, color = Region.x)) +
  geom_line() +
  labs(x = "Date", y = "Total Price", color = "Region", title = "Total Price over Time by Region") +
  theme_minimal()
```

### 19. Creating a scatterplot

```{r}
customer_sales_enriched |> 
  ggplot(mapping = aes(x = UnitPrice, y = Quantity, size = Profit)) +
  geom_point(alpha = 0.6, color = "lightblue") +  # Use alpha to make points semi-transparent for better visibility
  facet_wrap(~ Category) +   # Facet by Category
  labs(x = "Unit Price", y = "Quantity", size = "Profit", title = "Unit Price vs. Quantity by Category") +
  theme_minimal()
```

## Part 8: Advanced visualization

### 20. Faceting and Themes

Using `facet_wrap()` to create separate plots for each Region.

Applying a custom theme to enhance visual appeal.

```{r}
customer_sales_enriched |> 
  ggplot(mapping = aes(x = Date, y = TotalPrice)) +
  geom_line() +
  facet_wrap(~ Region.x) +
  theme_minimal()
```

### 21. Customize Scales and Guides

Customizing color scales and legends, using the `color` parameter.

Adding informative axis labels and a descriptive title using `labs()`.
```{r}
customer_sales_enriched |> 
  ggplot(mapping = aes(x = Date, y = TotalPrice, color = Category)) +
  geom_line() +
  facet_wrap(~ Region.x) +
  labs(x = "Date", y = "Total Price", color = "Category", title = "Total Price over Time by Category for each Region") +
  theme_minimal()
```

## Part 9: Additional Analysis with Joins

### 22. Identify unmatched records

After joins, identify any `OrderID` s without matching `CustomerID` s or `ProductID` s.

To do so, we use `filter()` with two conditions and `OR` statement between.

```{r}
customer_sales_enriched |> 
  filter(is.na(CustomerID) | is.na(ProductID))
```

> There are 3 rows with missing `CustomerID`

1. Possible reasons for missing `CustomerID`:
    1. Failure to capture customer information upon sale
    2. Customer purchasing online without registering a personal account
    3. Data handling mistakes, wrong / missing data entry
2. Potential implications:
    1. Missing information on customers - less information to base decisions on
    2. Not being able do identify loyal customers - missing opportunity to reward / incentivize
    3. Losing the capability of personalizing and targeting promotions to the individual customer

### 23. Compare Different Joins

Join customer_sales and customer_info using:

1. `left_join`

```{r}
left <- left_join(customer_sales, customer_info, by = 'CustomerID')
```
2. `inner_join`
```{r}
inner <- inner_join(customer_sales, customer_info, by = 'CustomerID')
```
3. `right_join`

```{r}
right <- right_join(customer_sales, customer_info, by = 'CustomerID')
```
4. `full_join`
```{r}
full <- full_join(customer_sales, customer_info, by = 'CustomerID')
```

Now, lets compare the number of rows received:
```{r}
nrow(left)
nrow(inner)
nrow(right)
nrow(full)
```

Differences between joins. When is each join type appropriate:

Join type | Method of working | When appropriate?
--|-----|------
Left | Takes the available values in the left data frame and looks for matching values in the right | Appropriate when we want to take the left's values as a basis and look up matching values for it
Inner | Only takes values from the 2 data frames where there are matching pairs | Appropriate when we only want the intersection of the 2 data frames
Right | The inverse of left: it takes the available values in the right data frame and looks for matching values in the left | Appropriate when we want to take the right's values as a basis and look up matching values for it in the left
Full | Joins the 2 data frames regardless of matching pairs. The resulting data frame includes all of the 2 data frame's values | Appropriate when we want to combine the values from 2 tables without leaving out any of the values in the join


## Part 10: Deliverables

### Data clearning

Our data clearing process started with importing necessary libraries to `R`. These included `dplyr`, `tidyverse`, `lubridate` and `readr`.

Then, data was imported from csv files using the `read_csv` command.

Upon inspecting data, we realized the size of it and necessary steps to perform.

1. First, we handled the missing values in `customer_sales` csv file. We filtered the tibble using the `filter()` command, to only see rows with
either `Quantity`, `UnitPrice` or `TotalPrice` missing. We then went on and inferred the missing values, if only just a single column was missing.
There were 10 cases in which we had to remove a row, because we could not inferr values - two or more columns were missing data.
2. We then progressed to standardizing date formats in the same file. During step 1, we've notised there are some cases when January is written as `Jän` (Austrian), not the German `Jan`.
These were replaced by using the `gsub()` command.
After these have beencleaned, we had to correct some dates, which were incorrectly parsed, due to using `YY-MM-DD` format. Some dates were from 2030! The `mutate()` command was very handy here.
3. We then fixed data types from `dbl` into `integer` or `date` using `as.integer()` and `as.Date()` commands.
4. Next, we integrated all csv files using joins. More about it in the joins part.
5. We also reshaped the tibbles with `tidyr`, using the `pivot_wider()` and `pivot_longer()` commands. This allowed us to display the Total Price (`sum`) per product in each of the regions easily.

### Analysis & Key findings

To analyze the data, we aggregated them to calculate Total Quantity, Price and Average profit per Sales Region, per Category (page 12 and 13).
From our analysis the biggest revenue driver is Furniture in North region with $19500$ dollars of total price, with the biggest average profit as well (443 dollars).

In total, however, its the West region that brings the most revenue (page 13), with 30100 in total.

We later visualized the data with `ggplot2`, creating a bar chart, showing the Electronics bringing less revenue than Furniture

```{r, echo=FALSE}
customer_sales_enriched |>
group_by(Category) |>
summarise(TotalPrice = sum(TotalPrice)) |>
ggplot(mapping = aes(x = Category)) +
geom_bar(aes(weight = TotalPrice))
```

Also, we've included a TimeSeries plot, to show the Total Price over Time by Region:

```{r, echo=FALSE}
customer_sales_enriched |>
ggplot(mapping = aes(x = Date, y = TotalPrice, color = Region.x)) +
geom_line() +
labs(x = "Date", y = "Total Price", color = "Region", title = "Total Price over Time by Region") +
theme_minimal()
```

And a scatterplot to compare Unit Price to Quantity by Category

```{r, echo=FALSE}
customer_sales_enriched |>
ggplot(mapping = aes(x = UnitPrice, y = Quantity, size = Profit)) +
geom_point(alpha = 0.6, color = "lightblue") + # Use alpha to make points semi-transparent for better
facet_wrap(~ Category) + # Facet by Category
labs(x = "Unit Price", y = "Quantity", size = "Profit", title = "Unit Price vs. Quantity by Category") +
theme_minimal()
```

We've used some advanced visualization methods as well, to display multiple plots on the same screen, or customize color scales.