---
title: "Unit 4"
author: Maciej Kilijański
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Comparing groups: statistical tests

Loading and summarizing the data:
```{r}
seg.df <- read.csv("https://goo.gl/qw303p",stringsAsFactors = TRUE)
summary(seg.df)
```

## Descriptive statistics

> Descriptive statistics for a specific segment can be obtained by indexing

Mean income for segment `moving up`
```{r}
mean(seg.df$income[seg.df$Segment == "Moving up"])
```

Mean income for segment `moving up` who do not subscribe
```{r}
mean(seg.df$income[seg.df$Segment == "Moving up" & 
                   seg.df$subscribe=="subNo"])
```

More efficient for multiple groups:
`by(data, INDICES,FUN)`
by divides data into groups for each unique value in indices and applies FUN to each group

```{r}
by(seg.df$income, seg.df$Segment, mean)
by(seg.df$income, list(seg.df$Segment, seg.df$subscribe), mean)
```

Better alternative with structured output (data frame), that can be used for further operations

> note: needs a list for even a single factor

```{r}
aggregate(seg.df$income, list(seg.df$Segment), mean)
```

We will use this output to add another variable to our data set that contains the segment mean:

```{r}
seg.income.mean <- aggregate(seg.df$income, list(seg.df$Segment), mean)
seg.df$segIncome <- seg.income.mean[seg.df$Segment, 2]
head(seg.df)
```

```{r, message=FALSE, warning=FALSE}
library(car)
```
```{r}
some(seg.df)
```

How did this work?
```{r}
head(seg.df$Segment)
head(seg.income.mean[seg.df$Segment, ])
head(seg.income.mean[seg.df$Segment, 2])
```
Clear that variable
```{r}
seg.df$segIncome <- NULL
```

### Formula version

R provides a standard way to describe relationships among
variables through formula specification.


A formula uses the tilde (~) operator to separate response variables on the left
from explanatory variables on the right. We will use this in different applications
```{r}
aggregate(income ~ Segment, data=seg.df, mean)
```

### Descriptives for two-way groups


Cross -tabulation is a common task in marketing, e.g., separating customers
into groups according to two or more factors
> two-way data aggregation

```{r}
aggregate(income ~ Segment + ownHome, data=seg.df, mean)
aggregate(income ~ Segment + ownHome + subscribe, data=seg.df, mean)
```
> Resulting output can be assigned to a data frame object and indexed
```{r}
agg.data <- aggregate(income ~ Segment + ownHome, data=seg.df, mean)
agg.data[2, ]
agg.data[2, 3]
```

Count of factor level occurence by factor
can be obtained by using the table expression
`with()` evaluates the expression in a data environment
same as `table(seg.df$Segment, seg.df$ownHome)`

```{r}
with(seg.df, table(Segment, ownHome))
```
Suppose we want a breakdown of the number of kids in each household (kids)
by segment. Here kids will be treated as factor and not as a number
```{r}
with(seg.df, table(kids, Segment))
```
> To obtain the total number of kids in each segment:

* Alternative 1:
```{r}
aggregate(kids ~ Segment, data=seg.df, sum)
```
* Alternative 2:
```{r}
seg.tab <- with(seg.df, table(kids, Segment))
apply(seg.tab*0:7, 2, sum)
```
* Alternative 3:
```{r}
seg.tab <- with(seg.df, table(kids, Segment))
colSums(seg.tab*0:7)
```