---
title: "Unit 3 - Tidyverse"
author: "Maciej Kilijański"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Slides [here](./unit_3_tidyverse.pdf)

# CRISP-DM data science workflow

> Cross-industry standard process for data mining (CRISP-DM)

## Sample workflow
1. Business understanding - what is the business problem/need?
2. Data understanding - What data do we have vs. need? What are our sources? How is data structured? Where is is stored?
3. Data preparation
4. Modeling - What modeling techniques should we use?
    - What's the granularity of forecast?
    - What are we predicting (*probability/sales volumes/etc* - format)
    - What are we judging the model on? *Accuracy of the forecast*
5. Evaluation - Which model best meets the business objectives?
6. Deployment - How do stakeholders access the results?

# Tidyverse

> A collection of `R` packages

* Consistent Syntax and philosophy

During the subjects following packages will be used:

* **dplyr** - Data Manipulation
* **tydyR** - Daty Cleaning (Tidying)
* **ggplot2** - Visualization

> Code written using Tidyverse is more **readeable** and **reproducible**

```{r, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
```

## Differences in syntax

### Pipes `%>%`

Pipes allow for composing a sequence of function calls in a more readeable way. Look at the following:

Standard R syntax:
```{Standard R syntax}
print(head(iris))
```

Pipes syntax
```{Pipes syntax}
iris %>% head() %>% print()
```

## Excercise

Sample data
```{r}
data('mtcars')
```

### Data manipulation with dplyr

#### `filter()` Cars with more than 20 miles per gallon and 6 cylinders 
Standard `R`
```{r}
mtcars_filtered <- mtcars[mtcars[,1] >= 20 & mtcars[,2] == 6,]
mtcars_filtered
```

`R` with `dplyr`

```{r}
mtcars_filtered <- mtcars %>% filter(mpg > 20) %>% filter (cyl == 6)
mtcars_filtered
```

#### `select()` columns by names or conditions (`head(10)` = 10 top rows)

```{r}
mtcars %>% select(mpg, cyl, hp) %>% head(10)
```

#### Calculate horse power to cylinder ratio

```{r}
mtcars_mutated <- mtcars %>% mutate(hp_per_cyl = hp / cyl) %>% 
select (hp, cyl, hp_per_cyl) %>% head(10)
mtcars_mutated
```

#### `arrange()` sort rows by columns

```{r}
mtcars %>% arrange(desc(mpg)) %>% head(10) %>% print()
```

#### `summarize()` and `group_by()`

```{r}
mtcars %>% group_by(cyl) %>% 
summarize(mean_mpg = mean(mpg), count = n()) %>% print()
```

### Data Tidying with tidyR

Example messy data
```{r}
messy_data <- tibble(
  id = 1:4,
  treatment_a = c(3, 4, 6, 8),
  treatment_b = c(5, 6, 7, 9)
)
messy_data
```

#### `gather()`: Convert wide data to long format
```{r}
tidy_data <- messy_data %>% gather(key = "treatment", value = "response", treatment_a:treatment_b)
tidy_data
```
#### `spread()`: Convert long data to wide format
```{r}
wide_data <- tidy_data %>% spread(key = treatment, value = response)
wide_data
```
#### `separate()`: Split a single column into multiple columns
```{r}
separated_data <- tidy_data %>% separate(treatment, into = c("treatment", "type"), sep = "_")
separated_data
```
#### `unite()`: Combine multiple columns into one
```{r}
united_data <- separated_data %>% unite(treatment_type, treatment, type, sep = "_")
united_data
```

### Advanced dplyr Techniques

#### Joins: `left_join()`, `inner_join()`, `right_join()`, `full_join()`
```{r}
df1 <- tibble(id = 1:3, value1 = c("A", "B", "C"))
df1 %>% print()
df2 <- tibble(id = 2:4, value2 = c("D", "E", "F"))
df2 %>% print()
```

Includes all values from df1 (left table), excludes all not-matched rows of df2 (right table)
```{r}
left_join(df1, df2, by = "id")
```
Includes only rows with matching keys in both joined tables
```{r}
inner_join(df1, df2, by = "id")
```
Includes all values from df2 (righ table), excludes all not-matched rows of df1 (left table)
```{r}
right_join(df1, df2, by = "id")
```
Includes all values
```{r}
full_join(df1, df2, by = "id")
```
#### Window functions: `lead()`, `lag()`, `cumsum()`, etc.
```{r}
mtcars %>% mutate(lag_mpg = lag(mpg), cumsum_mpg = cumsum(mpg)) %>%
select(mpg, lag_mpg, cumsum_mpg) %>% head(10)
```