---
title: "Unit 3 - Tidyverse"
author: "Maciej Kilijański"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Slides [here](./unit_3_tidyverse.pdf)

# CRISP-DM data science workflow

> Cross-industry standard process for data mining (CRISP-DM)

## Sample workflow
1. Business understanding - what is the business problem/need?
2. Data understanding - What data do we have vs. need? What are our sources? How is data structured? Where is is stored?
3. Data preparation
4. Modeling - What modeling techniques should we use?
    - What's the granularity of forecast?
    - What are we predicting (*probability/sales volumes/etc* - format)
    - What are we judging the model on? *Accuracy of the forecast*
5. Evaluation - Which model best meets the business objectives?
6. Deployment - How do stakeholders access the results?

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
